---
title: "ARER workflow - index calculation"
author: Michael Cecil
output: html_notebook
---

This notebook performs the rainfall index and payout calculations for much of the manuscript.
It uses different spatiotemporal assumptions on the CPC and CHIRPS datasets.


## load required libraries

```{r}

library(dplyr)
library(here)
library(sf)
library(raster)
library(conflicted)
library(rlang)
library(scico)
library(ggpubr)
library(assertthat)
library(ggplot2)

conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")
options(max.print = 10000)
```


## Load pre-calculated monthly averages for CPC data, filter to TX CPC grids

```{r}

monthly_averages <- read.csv(here("data/outputs/monthly_averages_na_rm.csv"))

## adjust longitude in above to be negative in western hemisphere
monthly_averages$lon <- monthly_averages$lon - 360

## Load CPC grids as spatial object, merge monthly average columns to spatial object
grids <- st_read(here("data/rainfall_index_grids/official_RMA_RI_grid.shp"))
grids$lon <- rowMeans(cbind(grids$X_MIN, grids$X_MAX))
grids$lat <- rowMeans(cbind(grids$Y_MIN, grids$Y_MAX))

monthly_averages_merge <- merge(grids, monthly_averages,
                          by = c("lon", "lat"))

## filter to TX grid cells
year <- 2023
tx_grids <- list.files(here("data/PRF-RI_CHIRPS/"), pattern = paste0("CHIRPS_precip_TX_", year, "_"), full.names = T) %>%
  first() %>% read.csv() %>% pull(GRIDCODE) %>% unique()

```

## merge CPC data with single year of monthly CHIRPS averags (to consolidate data columns)

```{r}

monthly_averages_tx <- monthly_averages_merge %>%
  filter(GRIDCODE %in% tx_grids)

current_year <- 2023

chirps_files <- list.files(here("data/PRF-RI_CHIRPS/"), pattern = paste0("CHIRPS_precip_TX_", current_year, "_"))
if(length(chirps_files) == 11){
  chirps_data <- read.csv(here(paste0("data/PRF-RI_CHIRPS/", chirps_files[grepl("625.csv", chirps_files)])))
  ## rename CHIRPS precip column with interval
  new_name <- paste0("CHIRPS_precip_", current_year, "_625")
  chirps_data[[new_name]] <- chirps_data$CHIRPS_precip
  chirps_data$CHIRPS_precip <- NULL

  ## merge with CPC data, need to have CHIRPS data on left as it has more rows
  chirps_data <- merge(chirps_data,
                       monthly_averages_tx,
                       by = 'GRIDCODE')

  for(interval in 626:635){
    chirps_data_temp <- read.csv(here(paste0("data/PRF-RI_CHIRPS/", chirps_files[grepl(paste0(interval, ".csv"), chirps_files)])))
  ## rename CHIRPS precip column with interval
    new_name <- paste0("CHIRPS_precip_", current_year, "_", interval)
    chirps_data_temp[[new_name]] <- chirps_data_temp$CHIRPS_precip
    chirps_data_temp$CHIRPS_precip <- NULL

  ## merge with CPC data, need to have CHIRPS data on left as it has more rows
    chirps_data <- merge(chirps_data,
                         chirps_data_temp[, c('CHIRPS_LAT', 'CHIRPS_LON', new_name)],
                         by = c('CHIRPS_LAT', 'CHIRPS_LON'))

  }
}

```



## Iterate over other years in CHIRPS data fiels, and join them to "chirps_data"
```{r}

for(year in 1981:2023){
  print(year)
  if(year == current_year){
    next
  }
  chirps_files <- list.files(here("data/PRF-RI_CHIRPS/"), pattern = paste0("CHIRPS_precip_TX_", year, "_"))
 
  if(length(chirps_files) != 11){
    print(paste0('incorrect # of files ', year))
    break
   }
  for(interval in 625:635){
    chirps_data_temp <- read.csv(here(paste0("data/PRF-RI_CHIRPS/", chirps_files[grepl(paste0(interval, ".csv"), chirps_files)])))
  ## rename CHIRPS precip column with interval
    new_name <- paste0("CHIRPS_precip_", year, "_", interval)
    chirps_data_temp[[new_name]] <- chirps_data_temp$CHIRPS_precip
    chirps_data_temp$CHIRPS_precip <- NULL   
  
  ## join data to existing chirps_data df
    chirps_data <- merge(chirps_data,
                         chirps_data_temp[, c('CHIRPS_LAT', 'CHIRPS_LON', new_name)],
                         by = c('CHIRPS_LAT', 'CHIRPS_LON'))
                         
  }
  
}

```


## Calculate CHIRPS monthly index for each interval, starting in year 1981
## We calculate the CHIRPS index starting in 2012, because we need a 30 year baseline (1981-2010) and one year gap (2011)
## Also we need to recalculate CPC index using a 1981 start year.

```{r}

## add CHIRPS index
for(year in 2012:2023){
  print(year)
  for(interval in 625:635){
  #  print(interval)
    current_year_precip <- chirps_data[[paste0("CHIRPS_precip_",
                                                    year,
                                                    "_",
                                                    interval) ]]
    baseline_precip_columns <- paste0("CHIRPS_precip_",
                                      1981:(year - 2),
                                      "_",
                                      interval)
    baseline_avg <- rowMeans(chirps_data[, baseline_precip_columns ])

    chirps_data[[paste0("CHIRPS_index_", year, "_", interval)]] <- round(current_year_precip/baseline_avg,  5)
  }
}

## add CPC index starting in 1981
for(year in 2012:2023){
  print(year)
  for(interval in 625:635){
    current_year_precip <- chirps_data[[paste0("precip_",
                                                    year,
                                                    ".",
                                                    interval) ]]
  #  print(current_year_precip)
    baseline_precip_columns <- paste0("precip_",
                                      1981:(year - 2),
                                      ".",
                                      interval)
  #  print(head(chirps_data[, baseline_precip_columns ]))
    baseline_avg <- rowMeans(chirps_data[, baseline_precip_columns ],
                             na.rm = T)
  #  print(baseline_avg)

    chirps_data[[paste0("cpc_index_1981_base_", year, "_", interval)]] <- round(current_year_precip/baseline_avg,  5)
  }
}

## save chirps_data file which contains all CPC and CHIRPS indices
save(chirps_data, file = paste0(here("data/outputs/"), "chirps_data_TX_ri.rda"))

```

## Calculate a "coarse" version of the CHIRPS index

```{r}

## calculate CHIRPS (coarse) precip and index values 

## coarsen raw precip
chirps_columns <- grep("CHIRPS_precip_", colnames(chirps_data), value = TRUE)

for(name in chirps_columns){
  print(name)
  chirps_data <- chirps_data %>%
    group_by(GRIDCODE) %>%
    mutate(!!paste0(name, "_coarse") := mean(!!sym(name), na.rm = TRUE)) %>%
    ungroup()
}

## update column names so that they end in year and interval
# Rename dataframe columns 
colnames(chirps_data) <- gsub(
  pattern = "^(CHIRPS_precip)_(\\d{4})_(\\d{3})_(coarse)$", 
  replacement = "\\1_\\4_\\2_\\3", 
  x = colnames(chirps_data)
)

## recalculate RI values for CHIRPS-coarse
## add CHIRPS (coarse) index
for(year in 2012:2023){
  print(year)
  for(interval in 625:635){
    current_year_precip <- chirps_data[[paste0("CHIRPS_precip_coarse_",
                                                    year,
                                                    "_",
                                                    interval, "") ]]
    baseline_precip_columns <- paste0("CHIRPS_precip_coarse_",
                                      1981:(year - 2),
                                      "_",
                                      interval, "")
    baseline_avg <- rowMeans(chirps_data[, baseline_precip_columns ])

    chirps_data[[paste0("CHIRPS_index_coarse_", year, "_", interval)]] <- round(current_year_precip/baseline_avg,  5)
  }
}

save(chirps_data, file = paste0(here("data/outputs/"), "chirps_data_TX_ri.rda"))

## view example
View(chirps_data %>% filter(GRIDCODE == 7329) %>% select(CHIRPS_precip_coarse_2017_630, CHIRPS_precip_2017_630, CHIRPS_index_coarse_2017_630, CHIRPS_index_2017_630))

print(chirps_data %>% filter(GRIDCODE == 7329) %>% select(CHIRPS_precip_coarse_2017_630, CHIRPS_precip_2017_630, CHIRPS_index_coarse_2017_630, CHIRPS_index_2017_630) %>% colMeans())
```
