---
title: "ARER workflow"
author: Michael Cecil
output: html_notebook
---

This notebook performs the rainfall index and payout calculations for much of the manuscript.
It uses different spatiotemporal assumptions on the CPC and CHIRPS datasets.


## load required libraries

```{r}

library(dplyr)
library(here)
library(sf)
library(raster)
library(conflicted)
library(rlang)
library(scico)
library(ggpubr)
library(assertthat)
library(ggplot2)

conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")
options(max.print = 10000)
```



## CDL proportions. Calculate county proportions based on CDL data (used in place of county to CPC grid proportions)
## outputs include a cropped and masked CDL raster for TX, 
## and a new file like "cpc_county_areas_cdl_2023.rda" that contains a column for CDL proportions
## this process assumes that the original CPC grid and county intersection list is correct (from "cpc-county-areas.csv")
## THIS TAKES A WHILE


```{r}

calc_cdl_prop <- function(year){
  
  # Load your cdl raster and tx counties
  r <- raster(here(paste0("data/cdl/raw/", year, "_30m_cdls/", year, "_30m_cdls.tif")))
  cdl_names <- read.csv(here("data/cdl/raw/CDL_codes_names_colors.csv"))
  tx_counties <- st_read(here("data/Tx_CntyBndry_Jurisdictional_TIGER/Tx_CntyBndry_Jurisdictional_TIGER.shp"))
  tx_counties$County.Code <- as.numeric(tx_counties$COUNTYID)
  sf_obj <- tx_counties
  
  # Reproject the tx counties to match the raster's CRS
  sf_obj_proj <- st_transform(sf_obj, crs = crs(r))
  
  # Crop and mask the raster to the sf object
  r_masked <- mask(crop(r, sf_obj_proj), sf_obj_proj)
  
  
  ## write cropped and masked cdl raster
  writeRaster(r_masked, file = here(paste0("data/outputs/cdl/cdl_tx_", year, ".tif")))
  
  # Get unique classes and their counts
  class_counts <- freq(r_masked, useNA = "no") %>%
    as.data.frame() %>%
    rename(class = value, count = count)
  
  
  # Calculate pixel area in square meters
  pixel_area <- res(r)[1] * res(r)[2]
  
  # Add area and percentage columns
  total_area <- sum(class_counts$count) * pixel_area
  class_counts <- class_counts %>%
    mutate(area_m2 = count * pixel_area,
           percentage = (area_m2 / total_area) * 100)
  
  # Display the results
  class_counts <- merge(class_counts, cdl_names,
                        by.x = "class",
                        by.y = "Codes")  %>% arrange(-count)
  print(class_counts %>% arrange(-count))
  
  write.csv(class_counts, here(paste0("data/outputs/cdl/class_counts_cdl_", year, ".csv")))
  
  
  
  grids <- st_read(here("data/rainfall_index_grids/official_RMA_RI_grid.shp"))
  grids$lon <- rowMeans(cbind(grids$X_MIN, grids$X_MAX))
  grids$lat <- rowMeans(cbind(grids$Y_MIN, grids$Y_MAX))
  
  grids_rpj <- st_transform(grids, st_crs(tx_counties))
  
  conflicts_prefer(raster::extract)
  
  cpc_county_areas <- read.csv(here("data/tx_county-areas/cpc-county-areas.csv"))
  
  cpc_county_areas$index <- 1:nrow(cpc_county_areas)
 # cpc_county_areas$CDL_area2 <- NA
  
  a <- sapply(1:nrow(cpc_county_areas), function(row){
    index <-cpc_county_areas$index[row] 
    print(index)
    cpc_grid <- cpc_county_areas$CPC.Grid.Code[row] 
    county_id <-cpc_county_areas$County.Code[row]  
    intersect_pred <- st_intersects(grids_rpj %>% filter(GRIDCODE == cpc_grid),
                                    tx_counties %>% filter(County.Code == county_id))
    # print(intersect_pred)
    if(length(intersect_pred[[1]]) == 0){
      cpc_county_areas$CDL_area2[index] <- 0
      return(0)
    }
    intersect <- st_intersection(grids_rpj %>% filter(GRIDCODE == cpc_grid),
                                 tx_counties %>% filter(County.Code == county_id))
    intersect_rpj <- st_transform(intersect, crs(r_masked))
    target_classes <- c(176, 37)
    pixel_vals <- extract(r_masked, intersect_rpj)[[1]]
    final_count <- sum(pixel_vals %in% target_classes, na.rm = T)
    # print('next')
    # print('next')
    # print(final_count)
    cpc_county_areas$CDL_area2[index] <- final_count
    return(final_count)
    
  })
  
  cdl_column_name <- paste0("cdl_area_", year) 
  
  cpc_county_areas$cdl_area <- a
  
  
  cpc_county_areas <- cpc_county_areas %>%
    group_by(County.Code) %>%
    mutate(Proportion_cdl = cdl_area/ sum(cdl_area))
  
  save(cpc_county_areas, file = here(paste0("data/outputs/cdl/cpc_county_areas_cdl_", year, ".rda")))
  
  print(paste0("cdl proportions for ", year, " completed"))
}

#calc_cdl_prop(2021)
calc_cdl_prop(2022)
#calc_cdl_prop(2023)


```
