---
title: "ARER Figures 2-1"
author: Michael Cecil
output: html_notebook
---


```{r}

library(dplyr)
library(here)
library(sf)
library(raster)
library(conflicted)
library(rlang)
library(scico)
library(ggpubr)
library(assertthat)
library(ggplot2)
library(exactextractr)
library(tidyr)
library(cowplot)
library(RColorBrewer)
library(grid)
library(gridExtra)
library(xtable)
library(patchwork)

conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")
options(max.print = 10000)


```


## code for exporting figures and tables to LaTex
```{r}
source(here("scripts/R/plotting_functions.R"))
```



## Figures tables for precipitation baselines, East and West TX
## Figure A.3, A.4


### East vs West Texas classification (counties)

```{r}

tx_counties <- st_read(here("data/Tx_CntyBndry_Jurisdictional_TIGER/Tx_CntyBndry_Jurisdictional_TIGER.shp"))
tx_counties <- st_set_crs(tx_counties, 4326)  # Example: EPSG 4326 (WGS84)
  
koppen_grid <- raster(here("data/koppen-climate-classification/koppen_ics/w001001.adf"))

tx_counties_rpj <- st_transform(tx_counties, crs = crs(koppen_grid))

## get unique classifications in TX
summary_stats <- exact_extract(koppen_grid, tx_counties_rpj %>% st_union(), function(values, coverage_fraction) {
    table(values)
}, progress = TRUE)

## 1 value is for "B" classifications (arid) 
## 2 value is for "C" classification (temperate)
reclass_df <- data.frame(c(7, 8, 15, 18, 22, 23),
                         c(1, 1, 2, 2, 2, 2))
koppen_grid_rcl <- reclassify(koppen_grid, rcl = as.matrix(reclass_df))

## extract values for TX counties
summary_stats_rcl <- exact_extract(
    koppen_grid_rcl,
    tx_counties_rpj,
    function(values, coverage_fraction) {
        tbl <- table(values)
        as.numeric(names(tbl)[which.max(tbl)])
    },
    progress = TRUE
)

tx_counties_rpj$region <- summary_stats_rcl
tx_counties$region <- summary_stats_rcl

east_tx_county_codes <- tx_counties %>% filter(region == 2) %>% pull(COUNTYID) %>% as.numeric()
west_tx_county_codes <- tx_counties %>% filter(region == 1) %>% pull(COUNTYID) %>% as.numeric()

plot(tx_counties_rpj['region'])

```


### East vs West Texas classification (CPC grids)

```{r}

## for CPC grids
grids <- st_read(here("data/rainfall_index_grids/official_RMA_RI_grid.shp"))
grids$lon <- rowMeans(cbind(grids$X_MIN, grids$X_MAX))
grids$lat <- rowMeans(cbind(grids$Y_MIN, grids$Y_MAX))

# monthly_averages_merge <- merge(grids, monthly_averages,
#                           by = c("lon", "lat"))



## filter to TX grid cells
year <- 2023
tx_grid_ids <- list.files(here("data/PRF-RI_CHIRPS/"), pattern = paste0("CHIRPS_precip_TX_", year, "_"), full.names = T) %>%
  first() %>% read.csv() %>% pull(GRIDCODE) %>% unique()

tx_grids <- grids %>% filter(GRIDCODE %in% tx_grid_ids)
plot(tx_grids)

tx_grids_rpj <- st_transform(tx_grids, crs(koppen_grid))

## get unique classifications in TX grids
summary_stats <- exact_extract(koppen_grid, tx_grids_rpj %>% st_union(), function(values, coverage_fraction) {
    table(values)
}, progress = TRUE)

reclass_df <- data.frame(c(7, 8, 15, 18, 19,  22, 23),
                         c(1, 1, 2, 2, 2, 2, 2))
koppen_grid_rcl <- reclassify(koppen_grid, rcl = as.matrix(reclass_df))

## extract values for TX grids
summary_stats_rcl <- exact_extract(
    koppen_grid_rcl,
    tx_grids_rpj,
    function(values, coverage_fraction) {
        tbl <- table(values)
        as.numeric(names(tbl)[which.max(tbl)])[1]
    },
    progress = TRUE
)

tx_grids_rpj$region <- summary_stats_rcl
tx_grids_rpj[tx_grids_rpj$GRIDCODE == 9719, "region"] <- 1
tx_grids_rpj[tx_grids_rpj$GRIDCODE %in%  c(7932, 8832, 9433), "region"] <- 2

plot(tx_grids_rpj["region"])


```


## Figure A.3
```{r}

tx_outline <- st_union(tx_counties) %>% st_transform(., crs = st_crs(tx_grids_rpj))

# Custom labels for the legend
legend_labels <- c("West Texas", "East Texas")

# Open a PNG graphics device
png(
  filename = here("data/outputs/figures/FigA3_tx_regions.png"),
  width = 6,    # Width in inches
  height = 6,  # Height in inches
  units = "in",
  res = 600         # Resolution in DPI
)

# Plot the sf object with customized legend
plot(
  tx_grids_rpj["region"],
  key.pos = 4,  # Position the legend (default is right of the plot)
  key.length = 1,  # Adjust the length of the legend
  key.width = 1,   # Adjust the width of the legend
  pal = colorRampPalette(c("blue", "green")),  # Define a color palette
  reset = FALSE   # Prevent resetting the plot window
)

plot(
  tx_outline,
  add = TRUE,      # Overlay on the existing plot
  border = "red",  # Set the border color to red
  lwd = 2          # Adjust the line width
)


# Add custom legend
legend(
  "topright",  # Position of the legend
  legend = legend_labels,  # Custom labels
  fill = colorRampPalette(c("blue", "green"))(2),  # Colors corresponding to labels
  title = "Region",
  bty = "n",  # Remove the legend box,
  cex = 0.8
)

  
  # Close the graphics device
dev.off()


```


### add region column to chirps data
```{r}

load(here("data/outputs/chirps_data_TX_ri.rda"))
## remove NA values
chirps_data <- chirps_data %>% filter(!is.na(cpc_index_1981_base_2012_625))

chirps_data <- merge(chirps_data, tx_grids_rpj %>% select(GRIDCODE, region),
                     by = "GRIDCODE") 

## save chirps data file with region for use in later plots
save(chirps_data, file = here("data/outputs/chirps_data_TX_region.rda"))

```


## Calculate baseline yearly precipitation
### Create new columns for yearly precipitation
```{r}
# Define suffixes to match (6 intervals, Jan-Feb, March-April)
# sum of 6 intervals is yearly precip
suffixes <- c("625", "627", "629", "631", "633", "635")

# Loop through the years and create new columns 
for (year in 1981:2023) {
  # Construct pattern to match columns for the specific year
  pattern <- paste0("CHIRPS_precip_", year, "_(", paste(suffixes, collapse = "|"), ")$")
  
  # Select relevant columns
  matching_cols <- grep(pattern, names(chirps_data), value = TRUE)
  
  # Sum the columns and add the new column
  chirps_data[[paste0("CHIRPS_precip_", year)]] <- rowSums(chirps_data[, matching_cols], na.rm = TRUE)
}

for (year in 1981:2023) {
  # Construct pattern to match columns for the specific year
  pattern <- paste0("^precip_", year, ".(", paste(suffixes, collapse = "|"), ")$")
  
  # Select relevant columns
  matching_cols <- grep(pattern, names(chirps_data), value = TRUE)
  
  # Sum the columns and add the new column
  chirps_data[[paste0("CPC_precip_", year)]] <- rowSums(chirps_data[, matching_cols], na.rm = TRUE)
}

```

### Some stats for response
### Mean precip during baseline (1981-2010), program (2012-23) years for E/W TX
```{r}

cpc_baseline_columns <- paste0("CPC_precip_", 1981:2010)
cpc_program_columns <- paste0("CPC_precip_", 2012:2023)
chirps_baseline_columns <- paste0("CHIRPS_precip_", 1981:2010)
chirps_program_columns <- paste0("CHIRPS_precip_", 2012:2023)

years <- 2012:2023
numbers <- 625:635

# Generate the vector of strings
cpc_program_ri_columns <- as.vector(outer(years, numbers, function(y, n) paste0("cpc_index_", y, "_", n)))
chirps_program_ri_columns <- as.vector(outer(years, numbers, function(y, n) paste0("CHIRPS_index_", y, "_", n)))


west_tx_mean_baseline_precip_cpc <- chirps_data %>% 
  filter(region == 1) %>% 
  select(all_of(cpc_baseline_columns)) %>% unlist() %>% mean(., na.rm = T)
west_tx_mean_baseline_precip_chirps <- chirps_data %>% 
  filter(region == 1) %>% 
  select(all_of(chirps_baseline_columns)) %>% unlist() %>% mean(., na.rm = T)
west_tx_mean_program_precip_cpc <- chirps_data %>% 
  filter(region == 1) %>% 
  select(all_of(cpc_program_columns)) %>% unlist() %>% mean(., na.rm = T)
west_tx_mean_program_precip_chirps <- chirps_data %>% 
  filter(region == 1) %>% 
  select(all_of(chirps_program_columns)) %>% unlist() %>% mean(., na.rm = T)
  
east_tx_mean_baseline_precip_cpc <- chirps_data %>% 
  filter(region == 2) %>% 
  select(all_of(cpc_baseline_columns)) %>% unlist() %>% mean(., na.rm = T)
east_tx_mean_baseline_precip_chirps <- chirps_data %>% 
  filter(region == 2) %>% 
  select(all_of(chirps_baseline_columns)) %>% unlist() %>% mean(., na.rm = T)
east_tx_mean_program_precip_cpc <- chirps_data %>% 
  filter(region == 2) %>% 
  select(all_of(cpc_program_columns)) %>% unlist() %>% mean(., na.rm = T)
east_tx_mean_program_precip_chirps <- chirps_data %>% 
  filter(region == 2) %>% 
  select(all_of(chirps_program_columns)) %>% unlist() %>% mean(., na.rm = T)


west_tx_mean_program_ri_cpc <- chirps_data %>% 
  filter(region == 1) %>% 
  select(all_of(cpc_program_ri_columns)) %>% unlist() %>% mean(., na.rm = T)
west_tx_mean_program_ri_chirps <- chirps_data %>% 
  filter(region == 1) %>% 
  select(all_of(chirps_program_ri_columns)) %>% unlist() %>% mean(., na.rm = T)
east_tx_mean_program_ri_cpc <- chirps_data %>% 
  filter(region == 2) %>% 
  select(all_of(cpc_program_ri_columns)) %>% unlist() %>% mean(., na.rm = T)
east_tx_mean_program_ri_chirps <- chirps_data %>% 
  filter(region == 2) %>% 
  select(all_of(chirps_program_ri_columns)) %>% unlist() %>% mean(., na.rm = T)




```





### Yearly Precip line plot, East TX
```{r}

## East TX

# Filter columns for years of interest for CHIRPS
yearly_chirps_cols <- grep("^CHIRPS_precip_\\d{4}$", names(chirps_data), value = TRUE)

# Reshape CHIRPS data to long format
chirps_long <- chirps_data %>%
  filter(region == 2) %>% 
  select(all_of(yearly_chirps_cols)) %>%
  pivot_longer(
    cols = everything(),
    names_to = "year",
    names_prefix = "CHIRPS_precip_",
    values_to = "precip"
  ) %>%
  mutate(year = as.numeric(year)) %>%
  group_by(year) %>%
  summarize(precip = mean(precip, na.rm = TRUE), .groups = "drop")  # Aggregate to one entry per year

# Filter columns for CPC dataset
yearly_cpc_cols <- grep("^CPC_precip_\\d{4}$", names(chirps_data), value = TRUE)

# Reshape CPC data to long format
cpc_long <- chirps_data %>%
  filter(region == 2) %>%
  select(all_of(yearly_cpc_cols)) %>%
  pivot_longer(
    cols = everything(),
    names_to = "year",
    names_prefix = "CPC_precip_",
    values_to = "precip"
  ) %>%
  mutate(year = as.numeric(year)) %>%
  group_by(year) %>%
  summarize(precip = mean(precip, na.rm = TRUE), .groups = "drop")  # Aggregate to one entry per year

# Combine both CHIRPS and CPC data
precip_long <- bind_rows(
  chirps_long %>% mutate(dataset = "CHIRPS"),
  cpc_long %>% mutate(dataset = "CPC")
)

# Calculate averages for the two periods
averages <- precip_long %>%
  filter(year %in% c(1981:2010, 2012:2023)) %>%
  group_by(period = case_when(
    year >= 1981 & year <= 2010 ~ "1981–2010",
    year >= 2012 & year <= 2023 ~ "2012–2023",
    TRUE ~ NA_character_
  ), dataset) %>%
  summarize(
    avg_precip = mean(precip, na.rm = TRUE),
    x_start = min(year),
    x_end = max(year),
    .groups = "drop"
  )

alpha_val <- 0.5



# Plot yearly precipitation for both CHIRPS and CPC data, with averages
p_east <- ggplot(precip_long, aes(x = year, y = precip, color = dataset)) +
  # CHIRPS 1981-2010
  geom_line(data = precip_long %>% filter(dataset == "CHIRPS" & year >= 1981 & year <= 2010),
            aes(color = "CHIRPS 1981–2010"), size = 1, alpha = alpha_val) +  
  # CPC 1981-2010
  geom_line(data = precip_long %>% filter(dataset == "CPC" & year >= 1981 & year <= 2010),
            aes(color = "CPC 1981–2010"), size = 1, alpha = alpha_val) +  
  # CHIRPS 2012-2023
  geom_line(data = precip_long %>% filter(dataset == "CHIRPS" & year >= 2012 & year <= 2023),
            aes(color = "CHIRPS 2012–2023"), size = 1, alpha = alpha_val) +  
  # CPC 2012-2023
  geom_line(data = precip_long %>% filter(dataset == "CPC" & year >= 2012 & year <= 2023),
            aes(color = "CPC 2012–2023"), size = 1, alpha = alpha_val) +  
  # Averages for each period with correct line colors
  geom_segment(
    data = averages %>% filter(period == "1981–2010" & dataset == "CPC"),
    aes(x = x_start, xend = x_end, y = avg_precip, yend = avg_precip, linetype = period, color = "CPC 1981–2010"),
    size = 1, alpha = alpha_val
  ) +
  geom_segment(
    data = averages %>% filter(period == "2012–2023" & dataset == "CPC"),
    aes(x = x_start, xend = x_end, y = avg_precip, yend = avg_precip, linetype = period, color = "CPC 2012–2023"),
    size = 1, alpha = alpha_val
  ) +
  geom_segment(
    data = averages %>% filter(period == "1981–2010" & dataset == "CHIRPS"),
    aes(x = x_start, xend = x_end, y = avg_precip, yend = avg_precip, linetype = period, color = "CHIRPS 1981–2010"),
    size = 1, alpha = alpha_val
  ) +
  geom_segment(
    data = averages %>% filter(period == "2012–2023" & dataset == "CHIRPS"),
    aes(x = x_start, xend = x_end, y = avg_precip, yend = avg_precip, linetype = period, color = "CHIRPS 2012–2023"),
    size = 1, alpha = alpha_val
  ) +
  scale_linetype_manual(
    values = c("1981–2010" = "dashed", "2012–2023" = "dashed"),
    name = "Period"
  ) +
  scale_color_manual(
    breaks = c( "CHIRPS 1981–2010", 
      "CPC 1981–2010", 
      "CHIRPS 2012–2023", 
      "CPC 2012–2023"),
    values = c(
       "orange", 
       "blue", 
      "orange4", 
       "blue4"
   #   "1981–2010" = "orange", 
    #  "2012–2023" = "orange4"
    ),
    name = ""
  ) +
  guides(color = guide_legend(order = 1), linetype = "none") +  # Remove the legend for Period (line type) and keep others
  labs(
   # title = "East TX Yearly Precipitation with Period Averages",
    x = "",
    y = ""
  ) +
  theme_minimal() +
  theme(text = element_text(size = 14)) + 
  theme(
    text = element_text(size = 14),
    legend.position = "bottom", 
    legend.box = "horizontal" 
  )   +
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  ) +
   annotate("text", x = 1980, y = max(precip_long$precip), 
           label = "East Texas", size = 6, hjust = 0, fontface = "bold") +
  clean_chart_clutter



```


### Yearly Precip line plot, West TX
```{r}


## West TX

# Filter columns for years of interest for CHIRPS
yearly_chirps_cols <- grep("^CHIRPS_precip_\\d{4}$", names(chirps_data), value = TRUE)

# Reshape CHIRPS data to long format
chirps_long <- chirps_data %>%
  filter(region == 1) %>% 
  select(all_of(yearly_chirps_cols)) %>%
  pivot_longer(
    cols = everything(),
    names_to = "year",
    names_prefix = "CHIRPS_precip_",
    values_to = "precip"
  ) %>%
  mutate(year = as.numeric(year)) %>%
  group_by(year) %>%
  summarize(precip = mean(precip, na.rm = TRUE), .groups = "drop")  # Aggregate to one entry per year

# Filter columns for CPC dataset
yearly_cpc_cols <- grep("^CPC_precip_\\d{4}$", names(chirps_data), value = TRUE)

# Reshape CPC data to long format
cpc_long <- chirps_data %>%
  filter(region == 1) %>%
  select(all_of(yearly_cpc_cols)) %>%
  pivot_longer(
    cols = everything(),
    names_to = "year",
    names_prefix = "CPC_precip_",
    values_to = "precip"
  ) %>%
  mutate(year = as.numeric(year)) %>%
  group_by(year) %>%
  summarize(precip = mean(precip, na.rm = TRUE), .groups = "drop")  # Aggregate to one entry per year

# Combine both CHIRPS and CPC data
precip_long <- bind_rows(
  chirps_long %>% mutate(dataset = "CHIRPS"),
  cpc_long %>% mutate(dataset = "CPC")
)

# Calculate averages for the two periods
averages <- precip_long %>%
  filter(year %in% c(1981:2010, 2012:2023)) %>%
  group_by(period = case_when(
    year >= 1981 & year <= 2010 ~ "1981–2010",
    year >= 2012 & year <= 2023 ~ "2012–2023",
    TRUE ~ NA_character_
  ), dataset) %>%
  summarize(
    avg_precip = mean(precip, na.rm = TRUE),
    x_start = min(year),
    x_end = max(year),
    .groups = "drop"
  )

alpha_val <- 0.5



# Plot yearly precipitation for both CHIRPS and CPC data, with averages
p_west <- ggplot(precip_long, aes(x = year, y = precip, color = dataset)) +
  # CHIRPS 1981-2010
  geom_line(data = precip_long %>% filter(dataset == "CHIRPS" & year >= 1981 & year <= 2010),
            aes(color = "CHIRPS 1981–2010"), size = 1, alpha = alpha_val) +  
  # CPC 1981-2010
  geom_line(data = precip_long %>% filter(dataset == "CPC" & year >= 1981 & year <= 2010),
            aes(color = "CPC 1981–2010"), size = 1, alpha = alpha_val) +  
  # CHIRPS 2012-2023
  geom_line(data = precip_long %>% filter(dataset == "CHIRPS" & year >= 2012 & year <= 2023),
            aes(color = "CHIRPS 2012–2023"), size = 1, alpha = alpha_val) +  
  # CPC 2012-2023
  geom_line(data = precip_long %>% filter(dataset == "CPC" & year >= 2012 & year <= 2023),
            aes(color = "CPC 2012–2023"), size = 1, alpha = alpha_val) +  
  # Averages for each period with correct line colors
  geom_segment(
    data = averages %>% filter(period == "1981–2010" & dataset == "CPC"),
    aes(x = x_start, xend = x_end, y = avg_precip, yend = avg_precip, linetype = period, color = "CPC 1981–2010"),
    size = 1, alpha = alpha_val
  ) +
  geom_segment(
    data = averages %>% filter(period == "2012–2023" & dataset == "CPC"),
    aes(x = x_start, xend = x_end, y = avg_precip, yend = avg_precip, linetype = period, color = "CPC 2012–2023"),
    size = 1, alpha = alpha_val
  ) +
  geom_segment(
    data = averages %>% filter(period == "1981–2010" & dataset == "CHIRPS"),
    aes(x = x_start, xend = x_end, y = avg_precip, yend = avg_precip, linetype = period, color = "CHIRPS 1981–2010"),
    size = 1, alpha = alpha_val
  ) +
  geom_segment(
    data = averages %>% filter(period == "2012–2023" & dataset == "CHIRPS"),
    aes(x = x_start, xend = x_end, y = avg_precip, yend = avg_precip, linetype = period, color = "CHIRPS 2012–2023"),
    size = 1, alpha = alpha_val
  ) +
  scale_linetype_manual(
    values = c("1981–2010" = "dashed", "2012–2023" = "dashed"),
    name = "Period"
  ) +
  scale_color_manual(
    breaks = c( "CHIRPS 1981–2010", 
      "CPC 1981–2010", 
      "CHIRPS 2012–2023", 
      "CPC 2012–2023"),
    values = c(
       "orange", 
       "blue", 
      "orange4", 
       "blue4"
   #   "1981–2010" = "orange", 
    #  "2012–2023" = "orange4"
    ),
    name = ""
  ) +
  guides(color = guide_legend(order = 1), linetype = "none") +  # Remove the legend for Period (line type) and keep others
  labs(
   # title = "West TX Yearly Precipitation with Period Averages",
    x = "",
    y = "Precip\n(mm)"
  ) +
  theme_minimal() +
  theme(text = element_text(size = 14)) + 
  theme(
    text = element_text(size = 14),
    legend.position = "bottom", 
    legend.box = "horizontal" 
  )    + # White background
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    axis.title.x = element_blank()
  ) +
   annotate("text", x = 1980, y = max(precip_long$precip), 
           label = "West Texas", size = 6, hjust = 0, fontface = "bold") +
  clean_chart_clutter
  

```



### Figure A.4
### Yearly Precip line plot, Combined

```{r}

combined_plot <- p_west + p_east +
  # Adjust layout to include the shared legend below
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom",  # Place the legend below the combined plot
    plot.title = element_text()
  ) 

p <- combined_plot + plot_annotation(tag_levels = "A") & 
  theme(
    plot.tag = element_text(face = "bold", size = 16),  # Bold tags
    plot.tag.position = c(0, 0.95)  # Adjust vertical position
  )

ggsave_ol(p, 
          "FigA4_west_east_tx_cpc_chirps_lines",  # Output file name
          "CPC and CHIRPS comparisons for historical baseline (1981-2010) and evaluation period (2012-2023)",
          height = 5,  # in inches
          width = 8   # in inches
)

```

